---
title: "Data Wrangling with R"
author: "Bernhard Piskernik"
output: 
  ioslides_presentation:
    logo: images/NOVOMATIC.png
    css: style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About me

Senior Data Scientist at Novomatic

Background in Humanities (Psychology)

## R {.build}

programming language for data analysis and statistics

it is free - Open Source-Software with GNU General Public License (GPL) 

it is also very popular in certain application areas (statistics, bioinformatics, ...) 

it is a dynamically typed interpreted language, typically used interactively

very extensible (>14k libraries at [CRAN](https://cran.r-project.org/))

interfaces to add functions in Fortran, C, C++, ...

## What makes R special?  {.build}

it is old (1992, but based on _S_ which was developed 1972)

it is explicitly build for data analysis and statistics (do not try to use it for general purpose programming)

it is 1-indexed

syntax WAS obnoxious before _tidyverse_

in combination with RStudio (IDE) very pleasant data wrangling and analysis experience

## The fundamental concept of R: vectors  {.build .smaller}

```{r a}
a <- rep(1,10)
```

```{r a_display, echo=FALSE}
a
```

```{r b}
b <- 1:10
```

```{r b_display, echo=FALSE}
b
```


```{r ab_plus}
a+b
```

```{r ab_calc}
a+2*b 
```

**Note: arithmetic operations act element-wise**

# DEMO
Building k-nearest neighbor classification from scratch

## Quick overview of kNN-classification (naive) {.build}

kNN is a non-parametric method for classification

* compute the distances from the test examples to all stored examples 

* get k-nearest neighbors per case

* conduct majority vote on class membership





