---
title: "kNN from Scratch"
output:
  html_document:
    df_print: paged
    selfcontained: true
---

# Preparations

load data wrangling package

```{r tidyvere}
library(tidyverse)
```


load data

```{r}
dat_raw <- read_csv("data.csv")
```


take a look at data

(Note: data is from [Vienna's land charge register](https://www.data.gv.at/katalog/dataset/stadt-wien_kaufpreissammlungliegenschaftenwien){target="_blank"} to which I have added [price clusters](http://www.bernhardlearns.com/2017/10/finding-viennas-house-price-regions.html){target="_blank"})

```{r}
dat_raw
```

quick visualization

```{r, message=FALSE}
## get map
library(ggmap)
vienna <- make_bbox(lon, lat, dat_raw) %>% 
  get_stamenmap(zoom = 13, maptype = c("toner-lite"))
ggmap(vienna, extent = "device") +
  coord_cartesian() +
  geom_point(data = dat_raw %>%
               filter(train==1),
             aes(x=lon, y=lat, color = as.factor(cluster)),
             alpha=.5) +
  geom_point(data = dat_raw %>%
               filter(train==0),
             aes(x=lon, y=lat),
             alpha=.5, shape=4) +
  theme(legend.position="none")
```



add id variable (= row number to data to identify individual cases)

```{r}
dat_id <- dat_raw %>%
  mutate(id = row_number())
```

select columns we will work with (id, lon, lat, cluster)

```{r}
dat_select <- dat_id %>%
  select(id, lon, lat, cluster)
```

separate training and test data

```{r}
## test data has no assigned 'cluster'
dat_test <- dat_select %>% 
  filter(is.na(cluster)) %>%
  ## remove cluster, because it is missing anyway
  select(-cluster)

## get train data by anti join with original data
dat_train <- dat_select %>%
  anti_join(dat_test, by="id")
```

# Get the nearest neighbors

cross training with test data (combine each case in training set with each case in the test set)

```{r}
dat_expand <- dat_train %>% 
  crossing(dat_test)
```

take a quick glance

```{r}
dat_expand
```

calculate euclidean distance (= square root of the sum of distance squares)

```{r}
dat_dist <- dat_expand %>%
  mutate(dist = sqrt((lat-lat1)^2 + (lon-lon1)^2))
```

get the k (=5) nearest cases per test set case

```{r}
dat_5nn <- dat_dist %>%
  ## segregate by id1
  group_by(id1) %>%
  ## get the 5 cases with smallest dist
  top_n(-5, dist)  %>%
  ## sort by id1
  arrange(id1)
```

get mayority vote

```{r}
dat_classified <- dat_5nn  %>%
  ## segregate by id1 and cluster
  group_by(id1, cluster) %>%
  ## get counts per resulting group
  count() %>%
  ## reduce to highest count
  top_n(1, n) %>%
  ## remove 'n'
  select(-n)
```

add result to test data

```{r}
dat_final <- dat_test %>%
  full_join(dat_classified, by=c("id"="id1"))
```


visually check result

```{r}
ggmap(vienna, extent = "device") +
  coord_cartesian() +
  geom_point(data = bind_rows(dat_train, dat_final),
             aes(x=lon, y=lat, color = as.factor(cluster)),
             alpha=.5) +
  theme(legend.position="none")
```


for reproducability

```{r}
sessionInfo()
```

