---
title: "kNN from Scratch"
output: html_notebook
---

# Preparations

load data wrangling package

```{r tidyvere}

```


load data ("data.csv")

```{r dat_raw}

```


take a look at data

(Note: data is from [Vienna's land charge register](https://www.data.gv.at/katalog/dataset/stadt-wien_kaufpreissammlungliegenschaftenwien){target="_blank"} to which I have added [price clusters](http://www.bernhardlearns.com/2017/10/finding-viennas-house-price-regions.html){target="_blank"})

```{r}

```

quick visualization

```{r, cache=TRUE, message=FALSE}
## get map
library(ggmap)
vienna <- make_bbox(lon, lat, dat_raw) %>% 
  get_stamenmap(zoom = 13, maptype = c("toner-lite"))
ggmap(vienna, extent = "device") +
  coord_cartesian() +
  geom_point(data = dat_raw %>%
               filter(train==1),
             aes(x=lon, y=lat, color = as.factor(cluster)),
             alpha=.5) +
  geom_point(data = dat_raw %>%
               filter(train==0),
             aes(x=lon, y=lat),
             alpha=.5, shape=4) +
  theme(legend.position="none")
```



add id variable (= row number to data to identify individual cases)

```{r dat_id}

```

select columns we will work with (id, lon, lat, cluster)

```{r dat_select}

```

separate training and test data

```{r dat_test__dat_train}
## test data has no assigned 'cluster'


## get train data by anti join with original data

```

# Get the nearest neighbors

cross training with test data (combine each case in training set with each case in the test set)

```{r dat_expand}

```

take a quick glance

```{r}

```

calculate euclidean distance (= square root of the sum of distance squares)

```{r dat_dist}

```

get the k (=5) nearest cases per test set case

```{r dat_5nn}

  ## segregate by id1
  
  ## get the 5 cases with smallest dist
  
  ## sort by id1
  
```

get mayority vote

```{r dat_classified}

  ## segregate by id1 and cluster
  
  ## get counts per resulting group
  
  ## reduce to highest count
  
  ## remove 'n'
  
```

add result to test data

```{r dat_final}

```


visually check result

```{r}

```


for reproducability

```{r}

```

